### 2014-02-11 Writing/plots/figureCode/MPA_figs.Rmd
Updated consMPA simulations

# Comparing MPAs to each other, to no management using figure plot code

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
```

```{r loadData, echo=FALSE, message=FALSE}
#sim=read.csv("~/Desktop/noMPAnotThresh_add_2013-12-09.csv")
# for emma's local copy
#sim=read.csv("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv")

# for synced online github data
x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/noMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL25vTVBBbm90VGhyZXNoX2FkZF8yMDEzLTEyLTA5LmNzdiIsImV4cGlyZXMiOjEzOTI2NDYxMDl9--46e45a3eb2faef7fc22b6124253c925bd2290de9")
sim = read.csv(textConnection(x))

fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/fishMPAnotThresh_add_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2Zpc2hNUEFub3RUaHJlc2hfYWRkXzIwMTMtMTItMDkuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE0NH0%3D--8eff4ee16d08efcf6b266c3e4e42adde7aeff969")
fishmpa = read.csv(textConnection(fish))

cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/nothresh/consMPAnotThresh_add_2014-01-23.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL25vdGhyZXNoL2NvbnNNUEFub3RUaHJlc2hfYWRkXzIwMTQtMDEtMjMuY3N2IiwiZXhwaXJlcyI6MTM5MjY0NjE2Nn0%3D--586d3cb31ac0f1fe3c49f48886e3796b5132bb77")
consmpa = read.csv(textConnection(cons))
```

```{r plotFunctions, echo=FALSE,message=FALSE}
plotfig <- function(df){
  cvals=unique(df$speed)
  hvals=unique(df$harvest)

  H=length(hvals)
  C=length(cvals)
  L=length(df$Equil.pop)

  ebm=array(,c(C,H))

  for(i in 1:L){
    hind=which(hvals==df$harvest[i])
    cind=which(cvals==df$speed[i])
    ebm[cind,hind]=df$Equil.pop[i]
  }

  all <- list(ebm = ebm, cvals = cvals, hvals = hvals, C = C, H = H)
  return(all)
}

plotParams <- function(lis){
  ebm <- lis$ebm
  cvals <- lis$cvals
  hvals <- lis$hvals
  C <- lis$C
  H <- lis$H
  # plotting parameters
  cex.lab=3
  cex.axis=2
  line=5
  lwd=5
  yaxs='i'
  xaxs='i'
  cex.lab=1.5
  cex.axis=1
  line=1
  oma=c(0,0,0,2)
  #breaks and colors
  lower=min(ebm)+1
  cuts=c(0,seq(lower,1538.649+.5,length.out=100))
  mycols=(sequential_hcl(length(cuts)-2,c.=c(0,0),l=c(30,100)))
  mycols=c('black',mycols)

  #axis labels
  numxlabs=11
  numylabs=3
  xdiff=.1
  ydiff=.2
  myaxes=list(arrows=FALSE,
              x=list(at=rev(seq(1,C,length.out=(cvals[C]-cvals[1])/xdiff+1)),labels=as.character(round(seq(cvals[1],cvals[C],by=xdiff),2))),
              y=list(at=rev(seq(1,H,length.out=(hvals[H]-hvals[1])/ydiff)),labels=as.character(round(seq(hvals[1],hvals[H],by=ydiff),2))),
              z=list(at=seq(0,1538.649,by=250),labels=as.character(seq(0,1538.649,by=250))))

  par(oma=oma)
  image.plot(cvals,hvals,ebm,breaks=cuts,col=mycols,
             xlab="Rate of environmental shift",ylab="Harvesting rate",
             cex.lab=cex.lab,cex.axis=cex.axis, 
             yaxs=yaxs,xaxs=xaxs,axes=TRUE,
             legend.shrink=1,legend.width=.1,zlim=range(cuts),
             axis.args=list(at=myaxes$z$at,labels=myaxes$z$labels,cex.axis=cex.axis),
             legend.args=list(text="Equilibrium biomass",cex=cex.lab,side=2,line=0.5,las=0),
             bigplot=c(.13,.85,.15,.95),smallplot=c(.91,.95,.15,.95),horizontal=FALSE)
  box()
}
```
Results of all simulations

```{r results}
par(mfrow=c(1,3))
noMang <- plotfig(sim); plotParams(noMang)
fish <- plotfig(fishmpa); plotParams(fish)
cons <- plotfig(consmpa); plotParams(cons)
```

Comparing no management to several small MPAs

```{r diff_fishnoManagement, echo=FALSE,message=FALSE}
noMang <- plotfig(sim)
fish <- plotfig(fishmpa)
combined <- noMang
combined$ebm = fish$ebm - noMang$ebm
plotParams(combined)
```

Comparing no management to few large MPAs

```{r diff_consnoManagement, message=FALSE,echo=FALSE}
noMang <- plotfig(sim)
cons <- plotfig(consmpa)
combined <- noMang
combined$ebm = cons$ebm - noMang$ebm
plotParams(combined)
```

Comparing the MPAs to each other (cons - fish)

```{r eachOther, message=FALSE, echo=FALSE, warning=FALSE}
combined <- cons
combined$ebm = cons$ebm - fish$ebm
plotParams(combined)
```