### 2014-02-12 Simluations/Aspatial_fast/Figures/MPA_diffs.Rmd
Reorganizing things

# Effect of MPAs

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
require(ggplot2)
require(ggthemes)
```
Have noticed that there's a section of parameter space in which the `Equil.pop` varies in the `consMPA` simulations. Why is that?

```{r loadData, echo=FALSE, message=FALSE}
# for synced online github data
   x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjEzfQ%3D%3D--9fbc7c170a93836981fa63448bf378fbbdb13905")
  sim = read.csv(textConnection(x))

  fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2013-12-09.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxMy0xMi0wOS5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjQzfQ%3D%3D--a432e4025f1b9042246c8a3d6b49af3464a03546")
  fishmpa = read.csv(textConnection(fish))

  cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-02-10.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMi0xMC5jc3YiLCJleHBpcmVzIjoxMzkyODI5NjYwfQ%3D%3D--00004e8d0186aa3a6f7f52bb264a406f147d1abf")
  consmpa = read.csv(textConnection(cons))

# combine all data into one dataframe
sim$management = rep("noManagement",nrow(sim))
consmpa$management = rep("consMPA", nrow(consmpa))
fishmpa$management = rep("fishMPA", nrow(fishmpa))

allData <- rbind(sim,consmpa,fishmpa)
```
Map equilibrium harvest as function of speed of climate velocity and harvest rate.

```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}
contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,3), drape=TRUE)

```

Looks like conservation MPAs do the best, but there are some weird fluctuations at higher harvest rates, why is that? Can take slices to look more closely

```{r variance, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.04 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

Trying the same plot but a higher rate of catch: $h = 0.10$ I get,

```{r variance_high, echo = FALSE, fig.width=10, fig.height=5}
subData <- subset(allData,harvest == 0.10 )
pd <- position_dodge(.1)
ggplot(subData, aes(x = speed, y = Equil.pop))+ facet_wrap(~management) + geom_line() + geom_ribbon(aes(ymin = Equil.pop - Equil.pop.se, ymax = Equil.pop + Equil.pop.se), alpha = 0.35) + theme_bw()
```

The relevant section of parameter space is shown below

```{r sub, echo=FALSE,fig.width=15, fig.height=5}
subData <- subset(allData,harvest > 0.05 & harvest < 0.13 & management == "consMPA" & speed < 0.3)
contourplot(Equil.pop ~ speed * harvest, data = subData, region = TRUE, drape = TRUE)
```

```{r sim, echo=FALSE}
setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
source("Parameters.R")
source("Functions.R")
require(zoo)
require(plyr)

# adding in subset of parameters, so instead of running Parameters_nothresh.R, am manually entering
    harvests = seq(0.05,0.13,0.001)
    speeds = seq(0,0.21,.001)

# build dataframes
  summaries <- data.frame(
		Equil.pop = rep(NA,length=length(speeds)*length(harvests)), 
		Equil.pop.sd = rep(NA, length=length(speeds)*length(harvests)), 
		Equil.pop.se = rep(NA, length=length(speeds)*length(harvests)), 
		Equil.harvest = rep(NA,length=length(speeds)*length(harvests)),
		Equil.harvest.sd = rep(NA,length=length(speeds)*length(harvests)), 
		Equil.harvest.se = rep(NA,length=length(speeds)*length(harvests)), 
		speed = rep(NA,length=length(speeds)*length(harvests)), 
		harvest = rep(NA,length=length(speeds)*length(harvests)), 
		thresh = rep(NA,length=length(speeds)*length(harvests)), 
		steps.harv = rep(NA,length=length(speeds)*length(harvests)),
		steps.move = rep(NA,length=length(speeds)*length(harvests)))

# index for row number
	rownumber <- matrix(seq(1:(length(harvests)*length(speeds))),ncol=length(speeds))

MPA = "cons"

source("Sim_noThresh.R")

```

After re-running the model for