### 2014-03-04 Simluations/Aspatial_fast/Figures/diagnose_consmpa.Rmd

# Figuring out the weird contour plot for `consmpa`

# Effect of MPAs

```{r required, echo=FALSE,message=FALSE}
library(fields)
library(lattice)
library(colorspace)
require(RCurl)
require(ggplot2)
require(ggthemes)
require(plyr)
require(reshape2)
require(RColorBrewer)
```
Have noticed that there's a section of parameter space in which the `Equil.pop` varies in the `consMPA` simulations. Why is that?

```{r loadData, echo=FALSE, message=FALSE}
# for synced online github data
x = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAnull_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQW51bGxfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzYwfQ%3D%3D--5fa73caaafffae7c95d930f61def04ee9b96930a")
sim = read.csv(textConnection(x))

fish = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAfish_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWZpc2hfMjAxNC0wMy0wMi5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzQ2fQ%3D%3D--9b6ceb0092173222543c95381dfbb0f42f0c884f")
fishmpa = read.csv(textConnection(fish))

cons = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/MPAcons_2014-03-01.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL01QQWNvbnNfMjAxNC0wMy0wMS5jc3YiLCJleHBpcmVzIjoxMzk0NDc3MzIwfQ%3D%3D--5f4f584256a4175eebe2ef5465e83c92ebe65cc0")
consmpa = read.csv(textConnection(cons))

thresh = getURL("https://raw.github.com/emfuller/MovingFish/master/Simluations/Aspatial_fast/Data/Thresh_2014-03-02.csv?token=3235371__eyJzY29wZSI6IlJhd0Jsb2I6ZW1mdWxsZXIvTW92aW5nRmlzaC9tYXN0ZXIvU2ltbHVhdGlvbnMvQXNwYXRpYWxfZmFzdC9EYXRhL1RocmVzaF8yMDE0LTAzLTAyLmNzdiIsImV4cGlyZXMiOjEzOTQ0ODIyMTV9--13d78257cda683d18eb8ecc71dcf0f4e6101b41d")
threshz = read.csv(textConnection(thresh))

# combine all data into one dataframe
sim$management = rep("No MPAs",nrow(sim))
consmpa$management = rep("Few Large MPAs", nrow(consmpa))
fishmpa$management = rep("Many Small MPAs", nrow(fishmpa))
threshz$management = rep("Thresholds", nrow(threshz))
allData <- rbind(consmpa,fishmpa)
```
Map equilibrium harvest as function of speed of climate velocity and harvest rate.

```{r Compare_Management,echo = FALSE, fig.width=10, fig.height=10}

# lattice plot settings

sb <- trellis.par.get("strip.background") 
sb[["col"]][1] <- "lightgray"
trellis.par.set("strip.background", sb)

ltheme <- canonical.theme(color = FALSE) ## in-built B&W theme 
ltheme$strip.background$col <- "transparent" ## change strip bg 
lattice.options(default.theme = ltheme) ## set as default

lw <- list(left.padding = list(x = 0, units = "inches")) 
lw$right.padding <- list(x = -0.1, units = "inches") 
lh <- list(bottom.padding = list(x = 0, units = "inches")) 
lh$top.padding <- list(x = -0.2, units = "inches")

lattice.options(layout.widths = lw, layout.heights = lh)

# mpa plots

contourplot(Equil.pop ~ speed * harvest | factor(management), data = allData, region=TRUE, layout = c(1,2), drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"))

#threshold plot

contourplot(Equil.pop ~ speed * -thresh | factor(management), data = threshz, region=TRUE, drape=TRUE,col.regions=rev(colorRampPalette(brewer.pal(9,"Greys"))(100)),colorkey=FALSE,labels=list(cex=.75,col="white"),scales=list( y=list(at=seq(-8,0,2),labels=round(rev(unique(threshz$thresh)/max(threshz$thresh)),3))),ylab=list(label="proportion below threshold"))

```

Looking at the weird variance in some of the `consmpa` simulations, try to target

```{r}
 ggplot(consmpa, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest)
```

Looks like it's in `h=0.08,0.09, 0.1, 0.12`

The relevant section of parameter space is shown below

```{r sub, echo=FALSE}
subData <- subset(consmpa,harvest > 0.07 & harvest < 0.13 & speed < 0.3 & speed > 0.1)
ggplot(subData, aes(x=speed, y = Equil.pop)) + geom_line() + facet_wrap(~harvest) + geom_point()
```
And it's in different places depending on the harvest.. So the simulations to look at are

+ `h = 0.08; s = 0.24-0.28`
+ `h = 0.09; s = 0.24-0.28`
+ `h = 0.10; s = 0.14-0.16`
+ `h = 0.12; s = 0.14-0.16`

The next step is to run a few of these simulations and save entire outputs (biomass over space/time). Will do `h = 0.08, s = 0.24-0.28`

```{r sim, echo=FALSE,fig.keep='none',message=FALSE,results='hide'}
setwd("/Users/efuller/Documents/Projects/Moving_fish/MovingFish/Simluations/Aspatial_fast")
source("Parameters.R")
source("Functions.R")
require(zoo)
require(plyr)
require(lattice)
require(dplyr)
require(reshape)
steps = 300   # setting number of steps over which to take equilibrium 
harvests = 0.08
speeds = seq(0.24,0.28,0.02)
MPA = "cons"
source("Figures/variable_steps.R")
```
What it looks like is that `h=0.08, s = 0.28` failed to converge. I finally killed it after 3821 generations.

```{r}
sims300 = sims
df300 = df
series300 = series
```
After re-running the model, can see that it replicates the pattern of fluctuating total biomass and huge error bars around `s = 0.26`
```{r final-plot,echo=FALSE,message=FALSE,results='hide'}
finals300 <- llply(series300,mean)
sd300 <- llply(series300,sd)
sd300 <- melt(sd300)
df_finals300 <- melt(finals300)
df_finals300$sd = sd300$value
names(df_finals300) <- c("totalPop","Speed","sd")
df_finals300$Speed = speeds
xyplot(totalPop ~ Speed, data=df_finals300,type='l',lwd=2,xlab="Speed", main = "300 steps")
ggplot(df_finals300,aes(x = Speed, y = totalPop)) + geom_errorbar(aes(ymin=totalPop - sd, ymax = totalPop + sd)) + geom_line() + geom_point()

```
And can see crazy fluctuations in `s = 0.26` that's not present in either of the other two simulations
```{r}
  xyplot(sums ~ time | speed, data = df300, type='l',xlab="time",ylab="Population biomass",lwd=2, main = "300 steps")
```

Re-running on `Sim_noThresh.R` for `s = 0.28` also doens't get convergence. So not sure why that simulation was able to run overnight.. Unless it takes just thousands of runs. But I think I could check that in the dataframe.

Switching to `s=0.26` issue, I think the issue may be that the `move` matrix is so small: it didn't run for many generations. So reaching back for 300 time steps involves a lot of the decline from when only harvest occurred, so it artificially inflates population biomass. So to check, look at the high `Equil.pop` and see how big the `move` matrix is.

```{r}
consmpa$steps.move[consmpa$harvest==0.08 & consmpa$speed==0.26]
consmpa$steps.move[consmpa$harvest==0.09 & consmpa$speed==0.26]
consmpa$steps.move[consmpa$harvest==0.1 & consmpa$speed==0.16]
consmpa$steps.move[consmpa$harvest==0.12 & consmpa$speed==0.16]
```
This still doesn't help me with the fact that `s = 0.28` now doesn't want to converge. But will leave that for now. Pretty sure that if I make the simulation run for a minimum of 500 generations. So need to target the parameter combinations that need to be re-run.

Look for differences within each speed simulation where the number of `steps.move` drops abruptly.

```{r}
ggplot(consmpa, aes(x=harvest, y = steps.move)) + geom_line() + geom_point() + facet_wrap(~speed)
```

Looks like all of `s=0.16` should be re-run, also all of `s=0.26`. This is because the number of generations dips below 400. The rest seems ok.